---
title: "Ed_X_intro_dataViz_R"
author: "Alfredo_Novoa"
date: "5/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intro To Data Viz ED X

https://rafalab.github.io/dsbook/introduction-to-data-visualization.html


- Plots of data easily communicate information that is difficult to extract from tables of raw values.
- Data visualization is a key component of exploratory data analysis (EDA), in which the properties of data are explored through visualization and summarization techniques.
- Data visualization can help discover biases, systematic errors, mistakes and other unexpected problems in data before those data are incorporated into potentially flawed analysis.
- This course covers the basics of data visualization and EDA in R using the ggplot2 package and motivating examples from world health, economics and infectious disease.

After completing Section 1, you will:

- Understand the importance of data visualization for communicating data-driven findings.
- Be able to use distributions to summarize data.
- Be able to use the average and the standard deviation to understand the normal distribution.
- Be able to assess how well a normal distribution fits the data using a quantile-quantile plot.
- Be able to interpret data from a boxplot.

There are 5 assignments that use the DataCamp platform for you to practice your coding skills.



```{r }
library(tidyverse)
library(dslabs)
library(ggplot2)

```

https://rafalab.github.io/dsbook/distributions.html

##### Key points
* The most basic statistical summary of a list of objects is its distribution.
* We will learn ways to visualize and analyze distributions in the upcoming videos.
* In some cases, data can be summarized by a two-number summary: the average and standard deviation. We will learn to use data visualization to determine when that is appropriate.

* Categorical data are variables that are defined by a small number of groups.
  + Ordinal categorical data have an inherent order to the categories (mild/medium/hot, for example).
  + Non-ordinal categorical data have no order to the categories.
* Numerical data take a variety of numeric values.
  + Continuous variables can take any value.
  + Discrete variables are limited to sets of specific values.
```{r}
data("heights")
names(heights)
head(heights) #what kind of variables are each

#lets find out what the unique values are 
#and how many of each there are
x <- heights$height
unique((x))
length(unique(x))

#we can also compute the freq of each unique value 
tab <- table(x)
tab

#now lets count the number of times a unique value appears only once
sum(table(x)==1)




```
### 1.2 Introduction to distributions 

This next section corresponds to the following sections:

* Textbook case study of [describing student heights](https://rafalab.github.io/dsbook/distributions.html#case-study-describing-student-heights)

* Textbook section on [the distribution function](https://rafalab.github.io/dsbook/distributions.html#distribution-function)
* Textbook section on the [cumulative distribution function](https://rafalab.github.io/dsbook/distributions.html#cdf-intro)
* Textbook section on [histograms](https://rafalab.github.io/dsbook/distributions.html#histograms)

* A distribution is a function or description that shows the possible values of a variable and how often those values occur.
For categorical variables, the distribution describes the proportions of each category.
* A *frequency* table is the simplest way to show a categorical distribution. Use `prop.table()` to convert a table of counts to a frequency table. *Barplots* display the distribution of categorical variables and are a way to visualize the information in frequency tables.
* For continuous numerical data, reporting the frequency of each unique entry is not an effective summary as many or most values are unique. Instead, a distribution function is required.
* The *cumulative distribution function* (CDF) is a function that reports the proportion of data below a value a for all values of a: F(a)=Pr(x≤a).
* The proportion of observations between any two values a and b can be computed from the CDF as F(b)−F(a).
* A histogram divides data into non-overlapping bins of the same size and plots the counts of number of values that fall in that interval.




```{r}

# make a table of category proportions
prop.table(table(heights$sex))
```

